### 2016-09-26 outline.rmd
adding ideas for talk.

# 1. Consumption and shopping
- largest supermarket chain in UK, famous for Clubcard starting in 1991; often cited as early example of Big Data Analytics
-

# 2. The web and the shopping basket
- online commerce -- transforming shopping;

>Tesco is the clear winner in the online grocery market, in fact it takes almost 50p of every £1 spent on food shopping on the internet. 
(https://econsultancy.com/blog/64841-how-seo-helps-tesco-to-dominate-the-online-grocery-market/)[https://econsultancy.com/blog/64841-how-seo-helps-tesco-to-dominate-the-online-grocery-market/]

# 3. From the rules-based list to a 'relevancy model'
How would I characterise this shift? Use it to emphasize why attention to the algorithm is not enough -- it is the model that counts
A little history of recommender systems -- uses Varian 1997 for a start. But has no real reference list. Not sure  
Also can use Schafer -- back to Association Rule Mining from early 1990s

# 4. Hack? Customers go straight to the checkout
people inhabit lists and work with them; no matter how they are made; cf ethnomethodology

# 5. Data: 52 weeks of data for each customer
- history of data accumulations;
- shift from data mining to personalization?

# 6. Gain: use good features to build models

# 7. Output 20 recommendations from 200,000 products
The problem of too many things to list -- this must be happening in many settings -- videos, social networks, products, etc.

# 8. Value: Balancing conversion vs spend
The value of a production

# 9. Metric: precision-at-K

# 10. Experiment: A/B allocation
Could use the Google Infrastructure of experiment or Facebook Flow stuff here -- 1 million models

# 11. Infrastructure: realtime engine serves up top 5
Only in certain settings are things realtime; or only in limited ways;
Realtime has its own economies and cultures

# 12. Similarity, substitution and complementarity
- interesting contrasts here between the relations of things in a list

# 13. Complementarity drives conversion 
- the problem of action