### 2017-10-05 reply_to_reviews.md
submitting rewritten version

Dear Evelyn
This is the re-written version of my earlier submission. I've pasting my response to the reviews below.

best wishes

Adrian Mackenzie

-----------------------------

Thanks very much to reviewers for their work on what must have seemed like a frustrating and elusive manuscript. The reviews have been a great help in re-writing the piece. The reviews were quite long, so I won't summarise all the changes I've made in response. There were many. The main focus in the re-writing of the manuscript was the overall argument.

Much of the call for clarification concerned the main argument. It has too many  threads. It was engaging in several different debates at once -- about recommender systems and their exploitative effects on  consumption practices, about recommender systems as examples of how big data practices take root in specific sectors, and recommender systems as symptom of some kind of more ontological transformation in sociotechnical collectives.  In the re-write, I've mainly sought to clarify the argument: grocery recommender systems are a mundane example of changes associated with big data; personalization is a powerful ordering technique in big data practice; but personalization cannot account for either the range of practices associated with big data or the broader associated transformations.  The case of  'conversion' of the Tesco grocery recommender system to a personal relevancy model helps make this argument: grocery shopping is a complex social negotiation, intimately entwined with social groups (family, household, etc). Data-mining to target grocery shoppers pre-dates by several decades recent growth in predictive analytics. A personalized recommender system, based even on post-demographic data, will struggle to effectively personalize grocery recommendations.

The manuscript has been rewritten from almost the ground up. It would be hard to list all the changes. Many of them respond to specific points in the reviews relating to relevant literature or problems in argumentation.   The basic idea of the rewrite was remove any thread of discussion that didn't relate to the main argument. A lot of the discussion of list practice and orderings was distracting so that has been reduced to a minimum. I wasn't able to provide the more substantial empirical engagement requested by Reviewer 2. That would entail a new research project.  The case study cannot make any claims to ethnographic thickness, so what the paper does to deepen its argumentative engagement around personalization in order to develop a more genealogical account based  on  research literature, widely used software and some examples of data from the Tesco API.

Main points from reviews:

1. Perceives **lack of cohesion and clarity in the argument;** re-frame opening to be less about commodities and supply chain, and more about problems of association, correlation, categorisation, etc.; establish a consistent role for the list; elaborate the methodology of reconstruction; clearly say what kind of transformation occurs in the transformation, and in the propensities; and what is the 'so what' of these clarifications and conceptualisation? 
2. Asks for more substantial empirical engagement, reorganization of the piece and **clarification of overall argument and purpose,** and specific section purposes; several dozen detailed points and calls for correction; 
3. Asks for clarification about the contribution of the piece; description of the case is vague in detail; look at Bucher and Barocas for state of the art; deal with disconnect between abstract and actual paper, especially around social order; **main argument is incomprehensible.**